---
title: "Phase 2: Agricultural Commodity Loss Variations across the Pacific Northwest Palouse Region, 1989-2015: Missing Data Examination"
author: "Erich Seamon"
date: "9/25/2018"
output:
  html_document: default
  pdf_document: default
params:
  output_dir: /dmine/code/git/dmine-paper1/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This analysis explores the relationships of agricultural commodity loss, at a county level, from 1989-2015, for the 26 county region of the Palouse, in Washington, Idaho, and Oregon. Here we explore the entire range of commodities and damage causes, identifying the top revenue loss commodities and their most pertinent damage causes - as indicated from the USDA's agricultural commodity loss insurance archive.

We perform several steps in Phase 2 to explore missing data, which is a continuation from Phase 1, steps 1-9.

**Step 10.**  Missing Data Examination.  In this step we examine missing data in relationship to all of our factors (year, county, damagecause) for each individual commodity we are examining (wheat, apples, cherries, and dry peas)
<br></br>
<br></br>


##Step 10: Missing Data Examination and filling in select zeros as NA

In Steps 7 and 8, we explore the missing data by county, damage cause, and year, and then fill in missing data for damage cause where commodity loss exists in a county.  For example, if wheat is grown in a county for a particular year, but NO wheat loss claims were submitted - we consider this a zero (there where zero claims - vs this data being missing).  We do this in order to ensure that we have as complete a dataset as possible.

```{r message=FALSE, warning=FALSE, echo=FALSE}

library(car)
library(RCurl)
library(lme4)
library(ez)
library(lattice)
library(ggplot2)
library(coefplot2)
library(broom)

#options(scipen=999)

#-load data
Southern_ID_sumloss <- read.csv(text=getURL
("https://raw.githubusercontent.com/erichseamon/dmine_anova/master/PNW_summary_all.csv"), 
header = TRUE)
Southern_ID_sumloss_all_sum  <- aggregate(loss ~ year + 
                                            damagecause + county + commodity,  
                                          Southern_ID_sumloss, sum)
Southern_ID_count_all_count  <- aggregate(count ~ year + 
                                            damagecause + county + commodity,  
                                          Southern_ID_sumloss, sum)
Southern_ID_sumloss_all_sum <- 
  Southern_ID_sumloss_all_sum[Southern_ID_sumloss_all_sum$loss >= 1, ]

#-Loading all WHEAT claims for the palouse from 1989-2015
palouse_sumloss <- read.csv(text=getURL
("https://raw.githubusercontent.com/erichseamon/dmine_anova/master/Palouse_summary_sumloss.csv"), 
header = TRUE)
palouse_counts <- read.csv(text=getURL
("https://raw.githubusercontent.com/erichseamon/dmine_anova/master/Palouse_summary_counts.csv"), 
header = TRUE)

#use a cube transformation on loss for WHEAT claims
Math.cbrt <- function(x) {
  sign(x) * abs(x)^(1/3)
}

Southern_ID_sumloss_all_sum$cube_loss <- Math.cbrt(Southern_ID_sumloss_all_sum$loss)
Southern_ID_count_all_count$cube_counts <- Math.cbrt(Southern_ID_count_all_count$count)

#--aggregate palouse
palouse_sumloss_aggregate <- aggregate(loss ~ damagecause + year + commodity + county, 
                                       palouse_sumloss, mean)

#-calculate cube loss
palouse_sumloss_aggregate$cube_loss <- Math.cbrt(palouse_sumloss_aggregate$loss)

#-remove zeros
palouse_sumloss_aggregate <- subset(palouse_sumloss_aggregate, loss > 0)

#-use a log transform 

palouse_sumloss_aggregate$log10_loss <- log10(palouse_sumloss_aggregate$loss)
palouse_sumloss_aggregate$log_loss <- log(palouse_sumloss_aggregate$loss)

#-inverse transform
palouse_sumloss_aggregate$inverse_loss <- 1/palouse_sumloss_aggregate$loss

#sq root transformation
palouse_sumloss_aggregate$sqroot_loss <- sqrt(palouse_sumloss_aggregate$loss)

#--scale and center
palouse_sumloss_aggregate$scaled_cube_loss <- scale(palouse_sumloss_aggregate$cube_loss)
palouse_sumloss_aggregate$scaled_inverse_loss <- 
  scale(palouse_sumloss_aggregate$inverse_loss, center = TRUE, scale = TRUE)



#----------


#--reduce years to 2001-2015
xxyear <- subset(palouse_sumloss_aggregate, year >= 2001)

#missing data review of data from 2001-2015 for all commodities and all damage causes
#ezDesign(xxyear, year, damagecause)
#ezDesign(xxyear, county, damagecause)
#ezDesign(xxyear, county, commodity)

#Now lets subset by the four main commodities of interest.
#--subset to four commodities - Barley, Wheat, Apples, and Dry Peas
xx <- subset(xxyear, commodity == "BARLEY" | commodity == "WHEAT" | 
               commodity == "APPLES" | commodity == "DRY PEAS" | commodity == "CHERRIES")

#--subset to a select set of damage causes - Drought, Heat, Hail, Frost, Freeze, Excessive Moisture, Cold Winter, Cold Weather, and Decline in Price
xxx <- subset(xx, damagecause == "Drought" | damagecause == "Heat" | 
                damagecause == "Hail" | damagecause == "Frost" | damagecause == "Freeze" | damagecause == "Excessive Moisture/Precip/Rain" | damagecause == "Cold Winter" | 
                damagecause == "Cold Wet Weather" | damagecause == "Decline in Price")


#examine missing data after narrowing commodities and damage causes to the most relevant
#ezDesign(xxx, year, damagecause)
#ezDesign(xxx, county, year)
#ezDesign(xxx, county, damagecause)
#ezDesign(xxx, year, commodity)
#ezDesign(xxx, county, commodity)

#palouse_sumloss_aggregate <- xxx

#now lets divide the data into four different model files that are commodity specific.
xxx_wheat <- subset(xxx, commodity == "WHEAT")
xxx_apples <- subset(xxx, commodity == "APPLES")
xxx_cherries <- subset(xxx, commodity == "CHERRIES")
xxx_drypeas <- subset(xxx, commodity == "DRY PEAS")
xxx_barley <- subset(xxx, commodity == "BARLEY")

palouse_sumloss_aggregate_apples <- xxx_apples
palouse_sumloss_aggregate_cherries <- xxx_cherries
palouse_sumloss_aggregate_drypeas <- xxx_drypeas
palouse_sumloss_aggregate_wheat <- xxx_wheat
palouse_sumloss_aggregate_barley <- xxx_barley

#-remove counties for each commodity file that have considerable missing data.  
#the reasoning is that those counties that have very little of the commodity in question
#may be inappropriate to fill in data as zero.  While there are sporadic commodity loss
#for these counties, there is a high liklihood that this commodity may NOT be grown
#for the missing years, vs just no commodity loss claims.  So we remove these select
#counties given the EzDesign data review of each commodity file

palouse_sumloss_aggregate_apples <- subset(palouse_sumloss_aggregate_apples, 
                                           county != "Columbia" & county != "Wasco")
palouse_sumloss_aggregate_apples$county <- 
  factor(palouse_sumloss_aggregate_apples$county)

palouse_sumloss_aggregate_apples <- 
  subset(palouse_sumloss_aggregate_apples, damagecause != "Drought")
palouse_sumloss_aggregate_apples$damagecause <- 
  factor(palouse_sumloss_aggregate_apples$damagecause)


palouse_sumloss_aggregate_wheat <- 
  subset(palouse_sumloss_aggregate_wheat, county != "Kootenai")
palouse_sumloss_aggregate_wheat$county <- 
  factor(palouse_sumloss_aggregate_wheat$county)

palouse_sumloss_aggregate_cherries <- 
  subset(palouse_sumloss_aggregate_cherries, county != "Adams" & county != "Union")
palouse_sumloss_aggregate_cherries$county <- 
  factor(palouse_sumloss_aggregate_cherries$county)

palouse_sumloss_aggregate_drypeas <- 
  subset(palouse_sumloss_aggregate_drypeas, county != "Douglas" 
         & county != "Gilliam" & county != "Adams")
palouse_sumloss_aggregate_drypeas$county <- 
  factor(palouse_sumloss_aggregate_drypeas$county)

palouse_sumloss_aggregate_barley <- 
  subset(palouse_sumloss_aggregate_barley, county != "Douglas" & county != "Garfield"
         & county != "Gilliam" & county != "Adams" & county != "Kootenai" & county != "Benton")
palouse_sumloss_aggregate_barley$county <- 
  factor(palouse_sumloss_aggregate_barley$county)

palouse_sumloss_aggregate_barley <- 
  subset(palouse_sumloss_aggregate_barley, damagecause != "Cold Winter")
palouse_sumloss_aggregate_barley$damagecause <- 
  factor(palouse_sumloss_aggregate_barley$damagecause)

#re-factor 

palouse_sumloss_aggregate_apples$damagecause <- factor(palouse_sumloss_aggregate_apples$damagecause)
palouse_sumloss_aggregate_wheat$damagecause <- factor(palouse_sumloss_aggregate_wheat$damagecause)
palouse_sumloss_aggregate_cherries$damagecause <- factor(palouse_sumloss_aggregate_cherries$damagecause)
palouse_sumloss_aggregate_drypeas$damagecause <- factor(palouse_sumloss_aggregate_drypeas$damagecause)
palouse_sumloss_aggregate_barley$damagecause <- factor(palouse_sumloss_aggregate_barley$damagecause)

palouse_sumloss_aggregate_wheat$commodity<- factor(palouse_sumloss_aggregate_wheat$commodity)
palouse_sumloss_aggregate_apples$commodity <- factor(palouse_sumloss_aggregate_apples$commodity)
palouse_sumloss_aggregate_barley$commodity <- factor(palouse_sumloss_aggregate_barley$commodity)
palouse_sumloss_aggregate_cherries$commodity <- factor(palouse_sumloss_aggregate_cherries$commodity)
palouse_sumloss_aggregate_drypeas$commodity <- factor(palouse_sumloss_aggregate_drypeas$commodity)

palouse_sumloss_aggregate_wheat$year <- factor(palouse_sumloss_aggregate_wheat$year)
palouse_sumloss_aggregate_apples$year <- factor(palouse_sumloss_aggregate_apples$year)
palouse_sumloss_aggregate_barley$year <- factor(palouse_sumloss_aggregate_barley$year)
palouse_sumloss_aggregate_cherries$year <- factor(palouse_sumloss_aggregate_cherries$year)
palouse_sumloss_aggregate_drypeas$year <- factor(palouse_sumloss_aggregate_drypeas$year)



#--create rows for missing data
alllevs_apples <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_apples[c("damagecause", "year", 
                                          "county", "commodity")], levels))
alllevs_wheat <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_wheat[c("damagecause", "year", 
                                         "county", "commodity")], levels))
alllevs_drypeas <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_drypeas[c("damagecause", "year", 
                                           "county", "commodity")], levels))
alllevs_cherries <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_cherries[c("damagecause", "year", 
                                            "county", "commodity")], levels))

alllevs_barley <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_barley[c("damagecause", "year", 
                                            "county", "commodity")], levels))

alllevs2_drypeas <- merge(palouse_sumloss_aggregate_drypeas, alllevs_drypeas, all.y=TRUE)
alllevs2_drypeas$loss[is.na(alllevs2_drypeas$loss)] <- 0
alllevs2_drypeas$cube_loss[is.na(alllevs2_drypeas$cube_loss)] <- 0
alllevs2_drypeas$log10_loss[is.na(alllevs2_drypeas$log10_loss)] <- 0


alllevs2_apples <- merge(palouse_sumloss_aggregate_apples, alllevs_apples, all.y=TRUE)
alllevs2_apples$loss[is.na(alllevs2_apples$loss)] <- 0
alllevs2_apples$cube_loss[is.na(alllevs2_apples$cube_loss)] <- 0
alllevs2_apples$log10_loss[is.na(alllevs2_apples$log10_loss)] <- 0


alllevs2_wheat <- merge(palouse_sumloss_aggregate_wheat, alllevs_wheat, all.y=TRUE)
alllevs2_wheat$loss[is.na(alllevs2_wheat$loss)] <- 0
alllevs2_wheat$cube_loss[is.na(alllevs2_wheat$cube_loss)] <- 0
alllevs2_wheat$log10_loss[is.na(alllevs2_wheat$log10_loss)] <- 0

alllevs2_cherries <- merge(palouse_sumloss_aggregate_cherries, alllevs_cherries, all.y=TRUE)
alllevs2_cherries$loss[is.na(alllevs2_cherries$loss)] <- 0
alllevs2_cherries$cube_loss[is.na(alllevs2_cherries$cube_loss)] <- 0
alllevs2_cherries$log10_loss[is.na(alllevs2_cherries$log10_loss)] <- 0

alllevs2_barley <- merge(palouse_sumloss_aggregate_barley, alllevs_barley, all.y=TRUE)
alllevs2_barley$loss[is.na(alllevs2_barley$loss)] <- 0
alllevs2_barley$cube_loss[is.na(alllevs2_barley$cube_loss)] <- 0
alllevs2_barley$log10_loss[is.na(alllevs2_barley$log10_loss)] <- 0

#testing the frequency of damagecauses for apples
#alllevs4_apples <- subset(alllevs2_apples, loss != 0.000000)
#alllevs4_apples <- aggregate(alllevs4_apples$loss, by = list(alllevs4_apples$damagecause), FUN = length)

#remove drought!  there is only one claim for drought for apples! 
#alllevs2_apples <- subset(alllevs2_apples, damagecause != "Drought")
#alllevs2_wheat <- subset(alllevs2_wheat, damagecause != "Drought")

#testing the frequency of damagecauses for wheat
#alllevs4_wheat <- subset(alllevs2_wheat, loss != 0.000000)
#alllevs4_wheat <- aggregate(alllevs4_wheat$loss, by = list(alllevs4_wheat$year, alllevs4_wheat$county), FUN = length)



```

<br></br>

###WHEAT Missing Data examination by damage cause, year, and county, using two dimensions

```{r message=FALSE, warning=FALSE, echo=FALSE}

#examine missing data after narrowing commodities and damage causes to the most relevant

ezDesign(palouse_sumloss_aggregate_wheat, year, damagecause)
ezDesign(palouse_sumloss_aggregate_wheat, year, county)
ezDesign(palouse_sumloss_aggregate_wheat, year, commodity)
ezDesign(palouse_sumloss_aggregate_wheat, county, commodity)
ezDesign(palouse_sumloss_aggregate_wheat, county, damagecause)


```


###APPLES Missing Data examination by damage cause, year, and county, using two dimensions

```{r message=FALSE, warning=FALSE, echo=FALSE}

#examine missing data after narrowing commodities and damage causes to the most relevant
ezDesign(palouse_sumloss_aggregate_apples, year, damagecause)
ezDesign(palouse_sumloss_aggregate_apples, year, county)
ezDesign(palouse_sumloss_aggregate_apples, year, commodity)
ezDesign(palouse_sumloss_aggregate_apples, county, commodity)
ezDesign(palouse_sumloss_aggregate_apples, county, damagecause)


```

###CHERRIES Missing Data examination by damage cause, year, and county, using two dimensions

```{r message=FALSE, warning=FALSE, echo=FALSE}

#examine missing data after narrowing commodities and damage causes to the most relevant
ezDesign(palouse_sumloss_aggregate_cherries, year, damagecause)
ezDesign(palouse_sumloss_aggregate_cherries, year, county)
ezDesign(palouse_sumloss_aggregate_cherries, year, commodity)
ezDesign(palouse_sumloss_aggregate_cherries, county, commodity)
ezDesign(palouse_sumloss_aggregate_cherries, county, damagecause)


```

###DRY PEAS Missing Data examination by damage cause, year, and county, using two dimensions

```{r message=FALSE, warning=FALSE, error = TRUE, echo=FALSE}

#examine missing data after narrowing commodities and damage causes to the most relevant
ezDesign(palouse_sumloss_aggregate_drypeas, year, damagecause)
ezDesign(palouse_sumloss_aggregate_drypeas, year, county)
ezDesign(palouse_sumloss_aggregate_drypeas, year, commodity)
ezDesign(palouse_sumloss_aggregate_drypeas, county, commodity)
ezDesign(palouse_sumloss_aggregate_drypeas, county, damagecause)

```


###BARLEY Missing Data examination by damage cause, year, and county, using two dimensions

```{r message=FALSE, warning=FALSE, error = TRUE, echo=FALSE}

#examine missing data after narrowing commodities and damage causes to the most relevant
ezDesign(palouse_sumloss_aggregate_barley, year, damagecause)
ezDesign(palouse_sumloss_aggregate_barley, year, county)
ezDesign(palouse_sumloss_aggregate_barley, year, commodity)
ezDesign(palouse_sumloss_aggregate_barley, county, commodity)
ezDesign(palouse_sumloss_aggregate_barley, county, damagecause)

```



```{r message=FALSE, warning=FALSE, echo=FALSE}

#--Prepare and Fill in missing data
palouse_sumloss_aggregate$year <- as.numeric(palouse_sumloss_aggregate$year)
palouse_sumloss_aggregate_apples$year <- as.numeric(palouse_sumloss_aggregate_apples$year)
palouse_sumloss_aggregate_wheat$year <- as.numeric(palouse_sumloss_aggregate_wheat$year)
palouse_sumloss_aggregate_drypeas$year <- as.numeric(palouse_sumloss_aggregate_drypeas$year)
palouse_sumloss_aggregate_cherries$year <- as.numeric(palouse_sumloss_aggregate_cherries$year)

#-turn year into character value
palouse_sumloss_aggregate_apples$year <- as.character(palouse_sumloss_aggregate_apples$year)
palouse_sumloss_aggregate_wheat$year <- as.character(palouse_sumloss_aggregate_wheat$year)
palouse_sumloss_aggregate_drypeas$year <- as.character(palouse_sumloss_aggregate_drypeas$year)
palouse_sumloss_aggregate_cherries$year <- as.character(palouse_sumloss_aggregate_cherries$year)

#-factor all group variables for each of the four commodities
palouse_sumloss_aggregate_apples$year <- factor(palouse_sumloss_aggregate_apples$year)
palouse_sumloss_aggregate_wheat$year <- factor(palouse_sumloss_aggregate_wheat$year)
palouse_sumloss_aggregate_drypeas$year <- factor(palouse_sumloss_aggregate_drypeas$year)
palouse_sumloss_aggregate_cherries$year <- factor(palouse_sumloss_aggregate_cherries$year)

palouse_sumloss_aggregate_apples$damagecause <- factor(palouse_sumloss_aggregate_apples$damagecause)
palouse_sumloss_aggregate_wheat$damagecause <- factor(palouse_sumloss_aggregate_wheat$damagecause)
palouse_sumloss_aggregate_drypeas$damagecause <- factor(palouse_sumloss_aggregate_drypeas$damagecause)
palouse_sumloss_aggregate_cherries$damagecause <- factor(palouse_sumloss_aggregate_cherries$damagecause)

palouse_sumloss_aggregate_apples$commodity <- factor(palouse_sumloss_aggregate_apples$commodity)
palouse_sumloss_aggregate_wheat$commodity <- factor(palouse_sumloss_aggregate_wheat$commodity)
palouse_sumloss_aggregate_drypeas$commodity <- factor(palouse_sumloss_aggregate_drypeas$commodity)
palouse_sumloss_aggregate_cherries$commodity <- factor(palouse_sumloss_aggregate_cherries$commodity)

palouse_sumloss_aggregate_apples$county <- factor(palouse_sumloss_aggregate_apples$county)
palouse_sumloss_aggregate_wheat$county <- factor(palouse_sumloss_aggregate_wheat$county)
palouse_sumloss_aggregate_drypeas$county <- factor(palouse_sumloss_aggregate_drypeas$county)
palouse_sumloss_aggregate_cherries$county <- factor(palouse_sumloss_aggregate_cherries$county)

#--create rows for missing data
alllevs_apples <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_apples[c("damagecause", "year", 
                                          "county", "commodity")], levels))
alllevs_wheat <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_wheat[c("damagecause", "year", 
                                         "county", "commodity")], levels))
alllevs_drypeas <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_drypeas[c("damagecause", "year", 
                                           "county", "commodity")], levels))
alllevs_cherries <- do.call(expand.grid, 
lapply(palouse_sumloss_aggregate_cherries[c("damagecause", "year", 
                                            "county", "commodity")], levels))

alllevs2_drypeas <- merge(palouse_sumloss_aggregate_drypeas, alllevs_drypeas, all.y=TRUE)
alllevs2_drypeas$loss[is.na(alllevs2_drypeas$loss)] <- 0
alllevs2_drypeas$cube_loss[is.na(alllevs2_drypeas$cube_loss)] <- 0
alllevs2_drypeas$log10_loss[is.na(alllevs2_drypeas$log10_loss)] <- 0


alllevs2_apples <- merge(palouse_sumloss_aggregate_apples, alllevs_apples, all.y=TRUE)
alllevs2_apples$loss[is.na(alllevs2_apples$loss)] <- 0
alllevs2_apples$cube_loss[is.na(alllevs2_apples$cube_loss)] <- 0
alllevs2_apples$log10_loss[is.na(alllevs2_apples$log10_loss)] <- 0


alllevs2_wheat <- merge(palouse_sumloss_aggregate_wheat, alllevs_wheat, all.y=TRUE)
alllevs2_wheat$loss[is.na(alllevs2_wheat$loss)] <- 0
alllevs2_wheat$cube_loss[is.na(alllevs2_wheat$cube_loss)] <- 0
alllevs2_wheat$log10_loss[is.na(alllevs2_wheat$log10_loss)] <- 0

alllevs2_cherries <- merge(palouse_sumloss_aggregate_cherries, alllevs_cherries, all.y=TRUE)
alllevs2_cherries$loss[is.na(alllevs2_cherries$loss)] <- 0
alllevs2_cherries$cube_loss[is.na(alllevs2_cherries$cube_loss)] <- 0
alllevs2_cherries$log10_loss[is.na(alllevs2_cherries$log10_loss)] <- 0

```